Increased perceptions of autonomy through choice fail to enhance motor skill retention

### 2021-06-16 docs/submission/manuscript.tex
init
update: submission and preprint dirs

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  jou]{apa7}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Increased perceptions of autonomy through choice fail to enhance motor skill retention},
  pdfauthor={Laura St.~Germain1, Allison Williams1, Noura Balbaa1, Andrew Poskus1, Olena Leshchyshen1, Keith R. Lohse2, \& Michael J. Carter1},
  pdflang={en-EN},
  pdfkeywords={Motor learning; Pre-registered; Self-controlled; Observation; Equivalence testing},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

@@

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{Autonomy-support does not benefit motor learning}
\keywords{Motor learning; Pre-registered; Self-controlled; Observation; Equivalence testing}
\usepackage{dblfloatfix}

\usepackage{csquotes}
\raggedbottom
\leftheader{\textbf{THIS PREPRINT HAS BEEN SUBMITTED FOR PEER-REVIEW}}
\rightheader{\textbf{THIS PREPRINT HAS BEEN SUBMITTED FOR PEER-REVIEW}}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

\addORCIDlink{Laura St. Germain}{0000-0002-5513-4183}; \addORCIDlink{Keith R. Lohse}{0000-0002-7643-3887}; \addORCIDlink{Michael J. Carter}{0000-0002-0675-4271}

We have outlined author contributions using CRediT (Contributor Roles Taxonomy - \url{https://casrai.org/credit/}).

The authors made the following contributions. Laura St.~Germain: Conceptualization, Data curation, Formal analysis, Investigation - Performed the experiment, Methodology, Project administration, Software - Task Programming, Validation, Visualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Allison Williams: Investigation - Performed the experiment, Writing - Original Draft Preparation, Writing - Review \& Editing; Noura Balbaa: Investigation - Performed the experiment, Writing - Original Draft Preparation, Writing - Review \& Editing; Andrew Poskus: Investigation - Performed the experiment, Writing - Review \& Editing; Olena Leshchyshen: Investigation - Performed the experiment, Writing - Review \& Editing; Keith R. Lohse: Conceptualization, Data curation, Formal analysis, Methodology, Validation, Writing - Original Draft Preparation, Writing - Review \& Editing; Michael J. Carter: Conceptualization, Data curation, Formal analysis, Funding acquisition, Methodology, Project administration, Resources, Software - Task Programming, Supervision, Validation, Visualization, Writing - Original Draft Preparation, Writing - Review \& Editing.

@@

\abstract{
There has been growing research interest in the effects that motivation plays in motor learning, and specifically how autonomy, competence, and social relatedness may directly benefit the learning process. Here, we present a preregistered manipulation of autonomy-support by providing learners with choice during the practice of a speed cup-stacking skill. One group was given control over when a video demonstration was provided and the viewing speed. A yoked control group received an identical demonstration schedule, but without choice (as their schedule was matched to a participant with choice). Critically, we addressed a gap in the literature by adding a yoked group who was explicitly told that they were being denied choice and that their schedule was chosen by another participant. We found no statistically significant learning differences between groups, despite finding evidence that providing choice increased perceived autonomy. Equivalence tests further showed that although the groups were not statistically equivalent, the effect size is likely too small to practically study the effects of autonomy-support through choice in most motor learning labs. These findings add to a growing body of research that questions a causal role of autonomy-support on motor learning, and the robustness of the so-called self-controlled learning advantage.
}

@@

\begin{table*}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:table1}Group specific instructions detailing the observation schedule that would be experienced during Acquisition.}

\begin{tabular}{ll}
\toprule
  & Acquisition instructions\\
\midrule
Self-Controlled & Before each trial, you will be asked whether you wish to watch a modeled demonstration of the task.\\
 & If you choose YES, you will then be asked whether you want to watch the video in real-time or in\\
 & slow-motion.\\
Traditional Yoked & Before each trial, you may or may not watch a modeled demonstration of the task based on a pre-\\
 & determined schedule. If you observe a model, it might be presented in real-time or in slow-motion\\
 & based on the pre-determined schedule.\\
Explicit Yoked & Before each trial, you may or may not watch a modeled demonstration of the task based on the\\
 & schedule another participant selected. If you observe a model, it might be presented in real-time or in\\
 & slow-motion based on what that participant selected.\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table*}

@@

Participants completed two data collection sessions separated by approximately 24 hours. Session 1 consisted of a pre-test and an acquisition phase. Session 2 consisted of a delayed retention test. At the start of each phase of the experiment, all participants received phase-specific instructions (see Data and code availability section). Group specific instructions were provided prior to the acquisition phase (see Table \ref{tab:table1}). The instructions appeared on a 22-inch computer monitor (1920x1080 resolution) positioned to the right of the participant. Participants followed along as the instructions were read aloud by the researcher.

@@

Our primary outcome measure was stacking time (i.e., response time) in seconds. Trials recorded as errors \((76/5250 = 1.45\%)\) during data collection were manually removed prior to data analysis. For each participant, pre-test and delayed retention trials were aggregated into one block of five trials and acquisition was aggregated into five blocks of five trials. Significance level was set to 0.05 for all statistical tests. Effect sizes for omnibus tests are reported using generalized eta squared \((\eta^2_{G})\) or eta squared \((\eta^2)\). Post hoc comparisons were conducted using Holm's correction. A Cook's distance of \(\geq 1\) was used to identify any influential cases.

@@

\begin{figure*}

{\centering \includegraphics[width=1\linewidth,height=1\textheight]{../../figs/fig1}

@@

\caption{Physical performance data of the experiment. The Self-Controlled group is shown in red with solid lines and/or circles, the Traditional Yoked group is shown in blue with dotted and/or triangles, and the Explicit Yoked group is shown in orange with dashed and/or squares. \textbf{A}. Trial-by-trial stacking time (s) data averaged across participants within each group (shaded area represents standard error). Splits between line segments denote the end of one phase and the start of the next. Pre-test (Trials 1 to 5) and Acquisition (Trials 6 to 30) occurred on Day 1 and Retention (Trials 31 to 35) occurred on Day 2, approximately 24-hours later. \textbf{B}. Mean stacking time (s) for each group was computed by averaging the data from \textbf{A} into 7 blocks of trials. Error bars represent 95\% confidence intervals. Splits between line segments denote the end of one phase and the start of the next. \textbf{C}. Mean stacking time (s) adjusted for pre-test for each group. Error bars represent 95\% confidence intervals. \textbf{D}. Scatterplot showing the relationship between each participant's mean stacking time (s) in pre-test and retention.}\label{fig:fig1}
\end{figure*}

\begin{table*}[t]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:table2}The mean frequencies for each acquisition block of requesting to watch the video demonstration and choosing the regular viewing speed.}

\begin{tabular}{llllll}
\toprule
  & Block 1 & Block 2 & Block 3 & Block 4 & Block 5\\
\midrule
Model viewing frequency & 32.6\% & 16.9\% & 12.0\% & 12.9\% & 9.3\%\\
 &  &  &  &  & \\
Regular speed frequency & 33.8\% & 38.1\% & 50.0\% & 34.4\% & 30.4\%\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table*}

@@

@@

To further probe our null effects (Harms \& Lakens, 2018) related to our main predictions, two equivalence tests (Self-Controlled versus Traditional Yoked and Traditional Yoked versus Explicit Yoked) on pre-test adjusted retention stacking times were carried out using the two one-sided test procedure (Lakens et al., 2018; Schuirmann, 1987). To establish our smallest effect size of interest (Lakens, 2021) we calculated the effect size our design had 33\% power to detect (Simonsohn, 2015), which was \(d_{s} = .31\). The Self-Controlled and Traditional Yoked groups were not statistically different, \(t(97.99) = -0.830\), \(p = 0.409\), and not statistically equivalent, \(t(97.99) = 0.720\), \(p = 0.237\), given equivalence bounds of \(-.355\) and \(.355\) (on a raw scale). Similarly, the Traditional Yoked and Explicit Yoked groups were not statistically different, \(t(97.99) = 0.262\), \(p = 0.794\), and not statistically equivalent, \(t(97.99) = -1.288\), \(p = 0.100\), given equivalence bounds of \(-.355\) and \(.355\) (on a raw scale).

\begin{figure*}

{\centering \includegraphics[width=0.85\linewidth,height=0.85\textheight]{../../figs/fig2}

\caption{Questionnaire data. Groups means for \textbf{A} perceived autonomy, \textbf{B} intrinsic motivation, and \textbf{C} perceived competence after pre-test, trial 5, and trial 25 on Day 1, and before retention on Day 2. The Self-Controlled group is shown in red circles, the Traditional Yoked group is shown in blue triangles, and the Explicit Yoked group is shown in orange squares. Dots represent individual participants in each group. Scores could range from 1 (Strongly disagree) to 7 (Strongly agree). Error bars represent 95\% confidence intervals.}\label{fig:fig2}
\end{figure*}

@@

It has been argued that self-controlled practice conditions are effective for motor learning because they are autonomy-supportive in nature (Wulf \& Lewthwaite, 2016). Yet, this claim has received, at best, modest support in the motor learning literature (e.g., McKay \& Ste-Marie, 2020a; but see Barros et al., 2019; Carter \& Ste-Marie, 2017b; Ste-Marie et al., 2013 for non-support). Here, we addressed a possible methodological limitation for this lack of support. That is, in previous self-controlled motor learning research, participants in the yoked group are not made aware that they have been denied opportunities to exercise choice over a practice variable such as feedback or watching a modeled demonstration. In the present experiment we introduced a novel yoked group that was explicitly told that the observation schedule---frequency and speed of video---they would receive during practice was actually one created by another participant in the experiment. Contrary to our prediction, the Explicit Yoked group did not report significantly lower perceived autonomy scores than the Traditional Yoked group, which was only informed their observation schedule was predetermined. In line with the ``OPTIMAL'' theory (Wulf \& Lewthwaite, 2016), the Self-Controlled group reported significantly higher perceived autonomy-support than both the Traditional and Explicit Yoked groups. However, this boost in perceived autonomy did not lead to enhanced motor performance or learning. These findings from a well-powered experiment are difficult to reconcile with a core pillar of the ``OPTIMAL'' theory of motor learning.

@@

In sum, we found that being able to control two aspects of one's observation schedule during practice was perceived as autonomy-supportive relative to not having this same control opportunity. However, making participants in an Explicit Yoked group aware they would be denied a control opportunity during practice did not decrease perceptions of autonomy relative to participants in the Traditional Yoked group. Furthermore, we did not find evidence than an autonomy-supportive manipulation had any direct causal effect on learning and we also failed to replicate the typical self-controlled learning advantage. It is worth noting that in some situations autonomy-support in and of itself might be a desired affective outcome (Ste-Marie et al., 2020). In such circumstance, providing learners with choice opportunities may be a useful motivationally-protective outcome. However, such a tool should be used judiciously as it is typically motor outcomes in the form of relatively permanent performance changes that is the true goal of any motor learning intervention (Ste-Marie et al., 2020). Ultimately, our results add to a growing body of evidence that questions whether autonomy-supportive manipulations directly affect learning. One explanation for this lack of replicability is that the \emph{``true''} effect of such manipulations are much smaller than previously estimated (McKay et al., 2021). Given the resources required (in terms of sample size) to reliably detect the tiny effects, we encourage motor learning scientists to invest their limited resources carefully---either studying practice conditions that have much bigger effects or devising ways to reduce variability and increase power through clever experimental design (McClelland, 2000).

@@

This work was supported by the Natural Sciences and Engineering Research Council (NSERC) of Canada (RGPIN-2018-05589; MJC). LSG was also supported by an NSERC PGS-D scholarship.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,floatsintext]{apa7}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Increased perceptions of autonomy through choice fail to enhance motor skill retention},
  pdfauthor={Laura St.~Germain1, Allison Williams1, Noura Balbaa1, Andrew Poskus1, Olena Leshchyshen1, Keith R. Lohse2, \& Michael J. Carter1},
  pdflang={en-EN},
  pdfkeywords={Motor learning; Pre-registered; Self-controlled; Observation; Equivalence testing},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

@@

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{Autonomy-support does not benefit motor learning}
\keywords{Motor learning; Pre-registered; Self-controlled; Observation; Equivalence testing\newline\indent Word count: 5055}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\usepackage{pdflscape}
\pagewiselinenumbers
\raggedbottom
\leftheader{\textbf{THIS PREPRINT HAS BEEN SUBMITTED FOR PEER-REVIEW}}
\rightheader{\textbf{THIS PREPRINT HAS BEEN SUBMITTED FOR PEER-REVIEW}}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

\addORCIDlink{Laura St. Germain}{0000-0002-5513-4183}

\addORCIDlink{Keith R. Lohse}{0000-0002-7643-3887}

\addORCIDlink{Michael J. Carter}{0000-0002-0675-4271}

Preregistration: \url{https://aspredicted.org/ze8cj.pdf}

Data, code, and materials: \url{https://github.com/cartermaclab/expt_explicit-yoked}

PsyArXiv preprint: \url{https://doi.org/10.31234/osf.io/s34fg}

@@

\abstract{
There has been growing research interest in the effects that motivation plays in motor learning, and specifically how autonomy, competence, and social relatedness may directly benefit the learning process. Here, we present a preregistered manipulation of autonomy-support by providing learners with choice during the practice of a speed cup-stacking task. One group was given control over when a video demonstration was provided and the viewing speed. A yoked control group received an identical demonstration schedule, but without choice (as their schedule was matched to a participant with choice). Critically, we addressed a gap in the literature by adding a yoked group who was explicitly told that they were being denied choice and that their schedule was chosen by another participant. We found no statistically significant learning differences between groups, despite finding evidence that providing choice increased perceived autonomy. Equivalence tests further showed that although the groups were not statistically equivalent, the effect size is likely too small to practically study the effects of autonomy-support through choice in most motor learning labs. These findings add to a growing body of research that questions a causal role of autonomy-support on motor learning, and the robustness of the so-called self-controlled learning advantage.

\textbf{\emph{Public Significance Statement}}\\
The present study suggests that although university-aged students may experience a boost in feelings of autonomy when given choice over features of their learning environment, this boost does not enhance learning of a new motor skill. Additionally, it highlights that coaches and practictioners should be encouraged to incorporate other and more effective practice variables into their training regimes.
}

@@

@@

Participants completed two data collection sessions separated by approximately 24 hours. Session 1 consisted of a pre-test and an acquisition phase. Session 2 consisted of a delayed retention test. At the start of each phase of the experiment, all participants received phase-specific instructions \hyperref[sec:sharing]{(see Data, materials, and code availability section)}. Group specific instructions were provided prior to the acquisition phase (see Table \ref{tab:table1}). The instructions appeared on a 22-inch computer monitor (1920x1080 resolution) positioned to the right of the participant. Participants followed along as the instructions were read aloud by the researcher. \medskip

\begin{table}

\caption{\label{tab:table1}Group specific components from the instructions detailing the observation schedule that would be experienced during Acquisition.}
\fontsize{11}{13}\selectfont
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{8em}>{\raggedright\arraybackslash}p{33em}}
\toprule
  & Acquisition instructions\\
\midrule
Self-Controlled & Before each trial, you will be asked whether you wish to watch a modeled demonstration of the task. If you choose YES, you will then be asked whether you want to watch the video in real-time or in slow-motion.\\
\addlinespace
Traditional Yoked & Before each trial, you may or may not watch a modeled demonstration of the task based on a pre-determined schedule. If you observe a model, it might be presented in real-time or in slow-motion based on the pre-determined schedule.\\
\addlinespace
Explicit Yoked & Before each trial, you may or may not watch a modeled demonstration of the task based on the schedule another participant selected. If you observe a model, it might be presented in real-time or in slow-motion based on what that participant selected.\\
\bottomrule
\end{tabular}
\end{table}

@@

Our primary outcome measure was stacking time (i.e., response time) in seconds. Trials recorded as errors \((76/5250 = 1.45\%)\) during data collection were manually removed prior to data analysis. For each participant, pre-test and delayed retention trials were aggregated into one block of five trials and acquisition was aggregated into five blocks of five trials. Significance level was set to 0.05 for all statistical tests. Effect sizes for omnibus tests are reported using generalized eta squared \((\eta^2_{G})\) or eta squared \((\eta^2)\). Post hoc comparisons were conducted using Holm's correction. A Cook's distance of \(\geq 1\) was used to identify any influential cases and none were identified.

@@

\begin{figure}

{\centering \includegraphics[height=0.51\textheight]{../../figs/fig1}

@@

\caption{Physical performance data of the experiment. The Self-Controlled group is shown in red with solid lines and/or circles, the Traditional Yoked group is shown in blue with dotted and/or triangles, and the Explicit Yoked group is shown in orange with dashed and/or squares. \textbf{A}. Trial-by-trial stacking time (s) data averaged across participants within each group (shaded area represents standard error). Splits between line segments denote the end of one phase and the start of the next. Pre-test (Trials 1 to 5) and Acquisition (Trials 6 to 30) occurred on Day 1 and Retention (Trials 31 to 35) occurred on Day 2, approximately 24-hours later. \textbf{B}. Mean stacking time (s) for each group was computed by averaging the data from \textbf{A} into 7 blocks of trials. Error bars represent 95\% confidence intervals. Splits between line segments denote the end of one phase and the start of the next. \textbf{C}. Mean stacking time (s) adjusted for pre-test for each group. Error bars represent 95\% confidence intervals. \textbf{D}. Scatterplot showing the relationship between each participant's mean stacking time (s) in pre-test and retention.}\label{fig:fig1}
\end{figure}

\begin{table}

\caption{\label{tab:table2}The mean frequencies for each Acquisition block of requesting to watch the video demonstration and of those requests, the mean frequency of choosing the regular viewing speed.}
\fontsize{11}{13}\selectfont
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{13em}>{\raggedright\arraybackslash}p{5em}>{\raggedright\arraybackslash}p{5em}>{\raggedright\arraybackslash}p{5em}>{\raggedright\arraybackslash}p{5em}l}
\toprule
  & Block 1 & Block 2 & Block 3 & Block 4 & Block 5\\
\midrule
Model viewing frequency & 32.6\% & 16.9\% & 12.0\%\% & 12.9\% & 9.3\%\\
\addlinespace
Regular speed frequency & 33.8\% & 38.1\% & 50.0\% & 34.4\% & 30.4\%\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{figure}

{\centering \includegraphics[height=0.72\textheight]{../../figs/fig2}

\caption{Questionnaire data. Groups means for \textbf{A} perceived autonomy, \textbf{B} intrinsic motivation, and \textbf{C} perceived competence after pre-test, trial 5, and trial 25 on Day 1, and before retention on Day 2. The Self-Controlled group is shown in red circles, the Traditional Yoked group is shown in blue triangles, and the Explicit Yoked group is shown in orange squares. Dots represent individual participants in each group. Scores could range from 1 (Strongly disagree) to 7 (Strongly agree). Error bars represent 95\% confidence intervals.}\label{fig:fig2}
\end{figure}

@@

To further probe the null effects (Harms \& Lakens, 2018) related to our main predictions, two equivalence tests (Self-Controlled versus Traditional Yoked and Traditional Yoked versus Explicit Yoked) on pre-test adjusted retention stacking times were carried out using the two one-sided test procedure (Lakens et al., 2018; Schuirmann, 1987). To establish our smallest effect size of interest (Lakens, 2021) we calculated the effect size our design had 33\% power to detect (Simonsohn, 2015), which was \(d_{s} = .31\). The Self-Controlled and Traditional Yoked groups were not statistically different, \(t(97.99) = -0.830\), \(p = 0.409\), and not statistically equivalent, \(t(97.99) = 0.720\), \(p = 0.237\), given equivalence bounds of \(-.355\) and \(.355\) (on a raw scale). Similarly, the Traditional Yoked and Explicit Yoked groups were not statistically different, \(t(97.99) = 0.262\), \(p = 0.794\), and not statistically equivalent, \(t(97.99) = -1.288\), \(p = 0.100\), given equivalence bounds of \(-.355\) and \(.355\) (on a raw scale).

@@

It has been argued that self-controlled practice conditions are effective for motor learning because they are autonomy-supportive in nature (Wulf \& Lewthwaite, 2016). Yet, this claim has received, at best, modest support in the motor learning literature (e.g., McKay \& Ste-Marie, 2020a; but see Barros et al., 2019; Carter \& Ste-Marie, 2017b; Ste-Marie et al., 2013 for non-support). Here, we addressed a possible methodological limitation for this lack of support. That is, in previous self-controlled motor learning research, participants in the yoked group are not made aware that they have been denied opportunities to exercise choice over a practice variable such as feedback or watching a modeled demonstration. In the present experiment we introduced a novel yoked group that was explicitly told that the observation schedule---frequency and speed of video---they would receive during practice was actually one created by another participant in the experiment. Contrary to our prediction, the Explicit Yoked group did not report significantly lower perceived autonomy scores than the Traditional Yoked group, which was only informed their observation schedule was predetermined. In line with the ``OPTIMAL'' theory (Wulf \& Lewthwaite, 2016), the Self-Controlled group reported significantly higher perceived autonomy-support than both the Traditional and Explicit Yoked groups. However, this boost in perceived autonomy did not lead to enhanced motor performance or learning. These findings from a preregistered experiment are difficult to reconcile with a core pillar of the ``OPTIMAL'' theory of motor learning.

@@

In sum, we found that being able to control two aspects of one's observation schedule during practice was perceived as autonomy-supportive relative to not having this same control opportunity. However, making participants in an Explicit Yoked group aware they would be denied a control opportunity during practice did not decrease perceptions of autonomy relative to participants in the Traditional Yoked group. Furthermore, we did not find evidence than an autonomy-supportive manipulation had any direct causal effect on learning and we also failed to replicate the typical self-controlled learning advantage. It is worth noting that in some situations autonomy-support in and of itself might be a desired affective outcome (Ste-Marie et al., 2020). In such circumstances, providing learners with choice opportunities may be a useful motivationally-protective outcome. However, such a tool should be used judiciously as it is typically motor outcomes in the form of relatively permanent performance changes that is the true goal of any motor learning intervention (Ste-Marie et al., 2020). Ultimately, our results add to a growing body of evidence that questions whether autonomy-supportive manipulations directly affect motor learning. One explanation for this lack of replicability is that the \emph{``true''} effect of such manipulations are much smaller than previously estimated (McKay et al., 2021). Given the resources required (in terms of sample size) to reliably detect these tiny effects, we encourage motor learning scientists to invest their limited resources carefully---either studying practice conditions that have much bigger effects or devising ways to reduce variability and increase power through clever experimental design (McClelland, 2000).

\vfill

\hypertarget{contributions-using-credit-httpscasrai.orgcredit}{%
\section{\texorpdfstring{Contributions using CrediT (\url{https://casrai.org/credit/})}{Contributions using CrediT (https://casrai.org/credit/)}}\label{contributions-using-credit-httpscasrai.orgcredit}}

Conceptualization (LSG, KRL, MJC); Data curation (LSG, KRL, MJC); Formal analysis (LSG, KRL, MJC); Funding acquisition (MJC); Investigation - Performed Experiment (LSG, AW, NB, AP, OL); Methodology (LSG, KRL, MJC); Project administration (LSG, MJC); Resources (MJC); Software - Task Programming (LSG, MJC); Supervision (MJC); Validation (LSG, KRL, MJC); Visualization (LSG, MJC); Writing - Original Draft Preparation (LSG, AW, NB, KRL, MJC); Writing - Review \& Editing (LSG, AW, NB, AP, OL, KRL, MJC)

\hypertarget{data-materials-and-code-availability}{%
\section{Data, materials, and code availability}\label{data-materials-and-code-availability}}

\label{sec:sharing}
All material, data, and scripts to reproduce our analyses and figures can be accessed here: \url{https://github.com/cartermaclab/expt_explicit-yoked}. The preregistration can be accessed here: \url{https://aspredicted.org/ze8cj.pdf}.

R {[}Version 4.1.0; R Core Team (2021){]} and the R-packages \emph{cowplot} {[}Version 1.1.1; Wilke (2020){]}, \emph{ggplot2} {[}Version 3.3.3; Wickham (2016){]}, \emph{jmv} {[}Version 1.2.23; Selker et al. (2020){]}, \emph{kableExtra} {[}Version 1.3.4; Zhu (2021){]}, \emph{papaja} {[}Version 0.1.0.9997; Aust and Barth (2020){]}, \emph{pwr} {[}Version 1.3.0; Champely (2020){]}, \emph{tidyverse} {[}Version 1.3.1; Wickham et al. (2019){]}, and \emph{TOSTER} {[}Version 0.4.0; Lakens (2017){]}.

@@

This work was supported by the Natural Sciences and Engineering Research Council (NSERC) of Canada (RGPIN-2018-05589; MJC). LSG was supported by an NSERC Postgraduate Scholarship. AW and AP were supported by NSERC Undergraduate Student Research Awards.
